Cohomology Theories, Triangulated Categories, and Applications\vspace{-1ex}

### 2019-06-03 Homological Algebra Notes/Hom Alg.tex
Update!

\usepackage{amsmath, amsfonts, amssymb, graphicx, color, fancyhdr, lipsum, scalerel, stackengine, mathrsfs, tikz-cd, mdframed, enumitem, framed, adjustbox, bm, upgreek, x	color}
\usepackage[framed,thmmarks]{ntheorem}
\usepackage[mathscr]{euscript}

@@

Let $\calA$ be a locally small category. Then $\mathbf{Funct}(\calA,\mathbf{Set})$ is the category of $\mathbf{Set}$-valued functors. Then there is a functor 
$\calA\to\mathbf{Funct}(\calA,\mathbf{Set})$ sending $A$ to $h_A=\Hom(A,-)$ (or $h^A=\Hom(-,A)$). Then the lemma says 
\begin{lem}[Yoneda]
	There exist bijections
	\[\Hom_\calA(A,A')\cong\mathbf{Nat}(h_{A'},h_A)\cong \mathbf{Nat}(h^A,h^{A'})\]
\end{lem}
\begin{cor}
	There is a fully faithful embedding of $\calA$ into $\mathbf{Funct}(\calA,\mathbf{Set})$.
\end{cor}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%              Homework                %%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

>>>>

\usepackage{amsmath, amsfonts, amssymb, graphicx, color, fancyhdr, lipsum, scalerel, stackengine, mathrsfs, tikz-cd, mdframed, enumitem, framed, adjustbox, bm, upgreek, x	color}
\usepackage[framed,thmmarks]{ntheorem}
\usepackage[mathscr]{euscript}
\usepackage{hyperref}

@@

Let $\calA$ be a locally small category. Then $\mathbf{Funct}(\calA,\mathbf{Set})$ is the category of $\mathbf{Set}$-valued functors. Then there is a functor 
$\calA\to\mathbf{Funct}(\calA,\mathbf{Set})$ sending $A$ to $h_A=\Hom(A,-)$ (or $h^A=\Hom(-,A)$). Then the lemma says 
\begin{lem}[Yoneda]
	There exist bijections
	\[\Hom_\calA(A,A')\cong\mathbf{Nat}(h_{A'},h_A)\cong \mathbf{Nat}(h^A,h^{A'})\]
\end{lem}
\begin{cor}
	There is a fully faithful embedding of $\calA$ into $\mathbf{Funct}(\calA,\mathbf{Set})$.
\end{cor}

\section{May 29, 2019}
\subsection{Motivation}
As some motivation for today, recall the problem on the first homework assignment where we were asked to 
say whether $\K(\Ch R)$ is abelian. The answer was no. Recall the 5-lemma (applied to triangles):

\begin{prop}
	If the following is a commutative diagram:
	\begin{center}
		\begin{tikzcd}
			A\ar[r]\ar[d,"\sim"] & B \ar[r]\ar[d,"\sim"] & C\ar[r]\ar[d,"h"] & \Sigma A\ar[d]\\
			A'\ar[r] & B'\ar[r] & C'\ar[r] & \Sigma A'
		\end{tikzcd}
	\end{center}
	is a morphism of triangles, then $h$ is an isomorphism.
\end{prop}
\begin{prf}
	Basically use Yoneda.
\end{prf}

\begin{rmk}
	In any triangulated category every monomorphism splits. To see this, take a triangle $A\xrightarrow{i}B\to C\to\Sigma A$.
	There is also an exact triangle $A\to A\oplus C\to C\xrightarrow{0}\Sigma A$ in $\K(R)$. Then use the five lemma and $B\cong A\oplus C$.
\end{rmk}
\begin{ex}
	We want to show that $\K(\bbZ)$ is not abelian. Let $C=\bbZ/4$ in degree zero and let $C'=\bbZ/2$ in degree zero. Let 
	$f:C\to C'$ be a morphism. If this category is abelian, then there exists a kernel of $f$, 
	so $\ker f\hookrightarrow C\xrightarrow{f}C'$ and furthermore this splits. 
	\[\bbZ/4\cong H(C)\not\cong H(C')\oplus H(\ker f)\] 
	but since $H(C')\cong\bbZ/2$, this is a contradiction.
\end{ex}
\subsection{Localization and the calculus of fractions}
Note that here set theory and size begins to be a problem, and this is discussed (e.g.) in Weibel. We 
won't delve into it much.
\begin{defn}
	Let $\calC$ be a category and $S$ a collection of morphisms in $\calC$. Then $S$ is called a multiplicative system of $\calC$ if 
	\begin{itemize}
		\item $S$ is closed under composition and contains all identity morphisms.
		\item (\textit{Ore Condition}) For any $s:Y\to X$ and $f:Z\to X$ are maps with $s\in S$,
		there are maps $t:W\to Z$ and $g:W\to Y$ with $t\in S$ such that 
		\begin{center}
			\begin{tikzcd}
				W\ar[r,"g"]\ar[d,"t"] & Y\ar[d,"s"]\\
				Z\ar[r,"f"] & X
			\end{tikzcd}
		\end{center}
		\item If $f,g:X\to Y$ are maps, then there exists a map $s\in S$ such that $s\circ f=s\circ g$ if 
		and only if there exists $t\in S$ such that $f\circ t=g\circ t$.
	\end{itemize}
\end{defn}

Then you get a new category $S^{-1}\calC$, the localizaton of $\calC$ at $S$. There is a functor $F:\calC\to S^{-1}\calC$ and it satisfies 
a universal property that if $G:\calC\to\calD$ is a functor for which $G(s)$ is an isomorphism for all $s\in S$, then $G$
factors through $S^{-1}\calC$.

We can do the usual ``roof construction'' on the $\Hom$ sets to complete the structure of this category.
\begin{thm}[Gabriel-Zisman]
	Let $S$ be a locally-small multiplicative system and let $S^{-1}\calC$ consist of the objects of $\calC$
	and whose morphisms are the roofs modulo the appropriate equivalence relation (see James' notes). This is a category and 
	satisfies the universal property.
\end{thm}
\begin{rmk}
	If $\calC$ is additive, then $S^{-1}\calC$ is additive.
\end{rmk}
\begin{rmk}
	The fact that all things in $S$ go to isomorphisms is a matter of playing with diagrams (roofs).
\end{rmk}

Using this, one can construct the derived category! Just localize $\K^\ast(R)$ a the quasi-isomorphisms.

\begin{defn}
	Let $\calA$ be an abelian category. Then $\calB$ is a Serre subcategory of $\calA$ if it is closed under
	\begin{itemize}
		\item subobjects
		\item quotients
		\item extensions
	\end{itemize}
\end{defn}
Then if we fix a subcategory $\calB$, let $S_\calB$ be the set of morphisms in an abelian category $\calA$ such that
both $\ker f$ and $\coker f$ lie in $\calB$, then 
\[\calA/\calB\cong S_\calB^{-1}\calA\]
is the quotient category.

\section{May 31, 2019}
Today I presented my talk on Brown representability. You can find the repository containing my notes for it \href{https://github.com/NicoCourts/Brown-Representability}{here}.

\section{June 3, 2019}
Today we're talking about spectral sequences! The primary references are Vakil's ``The Rising Sea'' (specifically section 1.7)
as well as Weiel and May's ``A primer on spectral sequences.''

\subsection{Crush Course in Spectral Sequences}
There are two usual treatments that are followed here: either filtered complexes or exact couples. In this lecture we will do neither.
Instead we will focus on a subcase of a filtered complex: double complexes. Recall
\begin{defn}
	A \textbf{double complex} $E_{\bullet\bullet}$ of a $\bbZ\times\bbZ$ indexed collection of objects $E_{pq}$ in some 
	abelian category, with differentials 
	\[d^v_{pq}:E_{pq}\to E_{p,q-1}\quad d^h_{p,q}:E_{p,q}\to E_{p-1,q}\]
	such that $d^v\circ d^v=d^h\circ d^h=0$ and 
	\[d^h\circ d^v+d^v\circ d^h=0\]
	(anticommutativity).
\end{defn}
So a diagram for a double complex (specifically a first-quadrant complex which we will be focusing on) is 
\begin{center}
\begin{tikzcd}
	later
\end{tikzcd}
\end{center}
\begin{defn}
	Recall that $\operatorname{Tot}(E_{\bullet\bullet})$ is the total complex where the $n^{th}$ piece is given by 
	\[\oplus_{i+j=n}E_{i,j}\]
	with differential $d=d^h+d^v$.
\end{defn}
The goal here is to compute information about the homology of this complex.
\subsection{Spectral Sequences}
\begin{defn}
	A \textbf{spectral sequence} is a sequence of \textbf{pages} $E_{p,q}^0,E_{p,q}^1,\dots$ for all $p$ and $q$ and all $r\ge 0$
	with differentials
	\[d_{p,q}^r:E_{p,q}^r\to E_{p-r,q+r-1}^r\]
	and fixed isomorphisms 
	\[E_{p,q}^{r+1}\cong H_{p,q}^r\]
	where $H_{p,q}^r$ is the homology on the $r^{th}$ page of $d^r$ at $E_{p,q}^r$.
\end{defn}

The zeroth page has vertical differentials, the first page has horizontal differentials. In general, the slope of the 
differential on the $r^{th}$ page is $\frac{r-1}{r}$.

\begin{rmk}
	For each $p,q$, there is an $r$ (large enough) such that any location stablizizes:
	that is, eventually the differential looks like 
	\[0\xrightarrow{d^r}E_{p,q}^r\xrightarrow{d^r}0\]
	and so $E_{p,q}^r\cong E_{p,q}^{r+1}$.

	This always happens when we are using a first quadrant complex, but whenever it happens, we will call 
	this stabilization point $E_{p,q}^\infty$. Notice where this happens is dependent on position, in general.
\end{rmk}

\begin{prop}
	$d^h$ induces $d^1$, the differential on the first page.
\end{prop}
\begin{prf}
	This amounts to showing that the obvious map is well-defined. :)
\end{prf}
\begin{prob}
	Show that $d^2$ is well-defined. This seems not completely obvious, but apparently it is a diagram chase away.
\end{prob}
\begin{rmk}
	Vakil apparently has an idea of ``$r$-closed $(p,q)$ strips'' that somehow obviate the need for such diagram chasing.
\end{rmk}

\subsection{Filtrations}
\begin{thm}
	$H_k(\operatorname{Tot}_\bullet(E_{\bullet\bullet}))$ is filtered by $E_{p,q}^\infty$ for $p+q=k$.

	More precisely, there is a filtration 
	\[0\subseteq F_0\subseteq \cdots\subseteq F_k\subseteq H_k(\operatorname{Tot}_\bullet(E_{\bullet\bullet}))\]
	such that $F_i/F_{i-1}\cong E_{i,k-i}^\infty$.
\end{thm}
\begin{rmk}
	In such a case, we say that $E_{p,q}^r$ \textbf{converges to} (or \textbf{abuts}) $H_{p+q}$ and write $E_{p,q}^r\Rightarrow H_{p+q}$.
\end{rmk}
\begin{rmk}
	We can ``flip'' across $y=x$ and use this spectral sequence. Notice the specific entries may be different.
\end{rmk}

\subsection{Applications}
We can use it to prove the snake lemma! Consider the diagram (which anticommutes):
\begin{center}
	\begin{tikzcd}
		0 & D\ar[d,"\alpha"]\ar[l] & E\ar[l]\ar[d,"-\beta"] & F\ar[l]\ar[d,"\gamma"] & 0\ar[l]\\
		0 & A\ar[l] & B\ar[l] & C\ar[l] & 0\ar[l]
	\end{tikzcd}
\end{center}

Then if we compute the horizontal direction first, we get $E_{pq}^1=E_{pq}^\infty=0$ so $H_k(\operatorname{Tot}_\bullet(E_{\bullet\bullet}))=0$ for all $k$.

Then consider the other direction (vertical first). From here you can extract the fact that you have an exact map between the kernels/cokernels 
and you get an isomorphism $d^2:E_{2,0}^2\to E_{0,1}^2$ on the second page, which basically gives you your connecting homomorphism.

\subsection{Further Reading and Applications}
The original context that these arose were in Leray's spectral sequence: Given an orientable fibration $F\hookrightarrow E\to B$, 
we get 
\[H^p(B,H^q(F))\Rightarrow H^\ast(E).\]

Another example is the Grothendieck spectral sequence. Given some ``nice'' functors $\calF$ and $\calG$, 
\[R^p\calF\circ R^q\calG(X)\Rightarrow R^{p+q}(\calF\circ \calG)(X)\]

Finally, (very recently!) Karim Adiprasito proved the ``g-conjecture.'' Along the way, he used certain spectral sequences 
of tensor products of chain complexes, and this motivated Sean's original research.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%              Homework                %%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

### 2019-06-03 Triangulated Categories/Tri Cat.tex
Update!

>>>>

\section{May 22, 2019}
\subsection{Example of the day}
This is a combination of some results of Artin, Van den Bergh, Yehutieli, Wu, Chan, and many others.

Suppose $A$ is a Noetherian algebra satisfying the following conditions:
\begin{itemize}
	\item $A$ is either connected graded or complete local
	\item $A$ is a finite module over it center.
\end{itemize}
Then there exists a dualizing complex $R$ over $(A,A)$, such that 
\begin{itemize}
	\item $R\Hom_A(k,R)\cong k$ and $R\Hom_{A^{op}}(k,R)\cong k$. This is sometimes called the \textit{pre-balanced condition.}
\end{itemize}

\begin{ex}
	Let $q$ be a root of unity. Let $A=k_q[[x_1,\cdots,x_n]]$ where we say any $x_i$ and $x_j$ $q$-commute, for $i\ne j$. Then this satisfies the hypotheses and $R$ exists.
	There is a method of constructing it via locak cohomology that is perhaps too much for what we have left in our course. We will, however, discuss local 
	cohomology and the $\chi$-condition next time.
\end{ex}

\subsection{Back to the future (Monday)}
We're going to tweak things a little.
\begin{thm}
	Let $A$ be a Noetherian local algebra. Let $X\in\Db_{fg}(\Amod)$ with $\projdim X<\infty$.
	Then there is a $Y\in\Db(\Amod)$ (here's where we weaken slightly) such that $\Hom_A(H^s(Y),k)\ne 0$ where 
	$s=\sup(Y)$. Then 
	\[\sup(R\Hom_A(X,Y))=\projdim X+\sup Y.\]
\end{thm}
\begin{prf}
	Last time we saw ``$\le$''. Now let $d=\projdim X$ and $s=\sup Y$. It suffices to show that
	\[H^{d+s}(R\Hom_A(X,Y))=H^{d+s}(\Hom_{\Ch(\Amod)}(P,Y))\ne 0\]
	where we take $P$ to be any (in this case minimal) free resolution of $X$. But if you consider the $d+s$ morphisms from $P$ to $Y$,
	there is only one map between nonzero things! So 
	\[\Hom_{\Ch(\Amod)}(P,Y)^{d+s}=\Hom_A(P^{-d},Y^s).\]

	We want to construct an $\tilde f\in\Hom_{\Ch}(P,Y)^{d+s}$ (notice it is already in the kernel) such that $\tilde f$ isn't in the image. To see this, 
	set $N=H^s(Y)=Y^s/\Im(\partial_Y^{s-1})\ne 0$ (we can assume $Y^{s+1}=0$). Then by the hypothesis, $\Hom_A(N,k)\ne 0$,
	but this is equal to 
	\[\Hom_A(N,A/J(A))=\Hom_k(A/J(A)\otimes N, A/J(A))\ne 0.\]

	This implies that $A/J(A)\otimes N\ne 0$ or $N/J(A)N\ne 0$. Notice that
	\[N/J(A)N=Y^s/(\Im(\partial_Y)+J(A)Y^s)\ne 0.\]

	Now since $N/J(A)N$ is a direct sum of simple modules, there is a nonzero morphism $f:P^{-d}\to N/J(A)N$. Since $P^{-d}$ is 
	(in particular) projective, we can lift to a morphism $\tilde f:P^{-d}\to Y^s$. Furthermore, $\Im(\tilde f)$ is not contained within
	$\Im(\partial_Y^{s-1})+J(A)Y^s$ (since it wasn't the zero map on $N/J(A)N$).

	Now consider the chain complex $\Hom_{\Ch(\Amod)}(P,Y)^\ast$ at the $d+s-1^{st}$ term. The differential maps from 
	$\Hom_A(P^-d,Y^{s-1})\oplus\Hom_A(P^{-d+1},Y^s)\to \Hom_A(P^{-d},Y^s)$ where 
	\[\partial(\alpha+\beta)=\partial_Y\circ\alpha+\beta\circ\partial_P\]

	But then by checking maps (and using that since $P$ is minimal, $\partial_P^{-d}$ maps into $J(A)P^{-d+1}$), we can 
	show that the image of any map under this differential must map into this object, so $\tilde f$ must not lie in the image, 
	so it is a nonzero element of the homology.
\end{prf}

\subsection{Tensor Product of Complexes}
Now let $X\in\Ch(\modA)$ and $Y\in\Ch(\Amod)$. Then $X\otimes_A Y\in\Ch(\Vectk)$ is
\[(X\otimes_A Y)^i=\bigoplus_{i+j=i}X^i\otimes Y^j\]
and the differential is given by the Koszul sign rule $\partial(x\otimes y)=\partial_X(x)\otimes y+(-1)^|x|x\otimes\partial_Y(y)$

The derived functor of $-\otimes_A-$ is 
\[X\otimes_A^LY\cong P_X\otimes_A y\cong X\otimes_A P_Y\cong P_x\otimes_AP_Y\]
where, for example, $A\otimes_A^LY=A\otimes_AY=Y$ and $X\otimes_A^LA=X$.

\begin{thm}
	Let $A$ be a Noetherian local algebra. Let $X\in\Db_{fg}(\Amod)$ with $\projdim X<\infty$. Let $Y\in\Db(\modA)$
	such that $\Hom_A(k,H^i(Y))\ne 0$ where $i=\inf Y$. Then 
	\[\inf(Y\otimes_A^LX)=-\projdim X+\inf Y\]
\end{thm}

\begin{lem}
	Let $A$ be Noetherian and $X\in\Db_{fg}(\Amod)$, $Y\in\Db(A^e\text{-}\mathbf{mod})$ and $Z\in \Dbperf(\Amod)$. Then 
	\[R\Hom_A(XY\otimes_A^LZ)\cong R\Hom_A(X,Y)\otimes_A^LZ\]
\end{lem}
\begin{prf}
	Let $F(-)=R\Hom_A(X,Y\otimes_A^L-)$ and $G(-)=R\Hom_A(X,Y)\otimes_A^L-$. We want to construct 
	a natural isomorphism between them. There is a map $\eta:F\to G$ that sends 
	\[f\otimes z\mapsto \{x\mapsto f(x)\otimes z\}.\]
	Then it only remains to check it is natural.

	The next step is to show that it holds (trivially!) for $Z=A$ and then since $A$ generates $\Dbperf(\Amod)$, we get that $G\cong F$ for all 
	things in $\Dbperf(\Amod)$.
\end{prf}

\section{May 24, 2019}
Had to miss this day.

\section{May 29, 2019}

\subsection{Example of the Day}
Let $A$ be a Noetherian algebra that is a finitely generated module over its center. Furthermore, suppose that 
$\calZ(A)\cong k[x_1,\cdots,x_n]/I$. Then $A$ has a dualizing complex $R$ where $R=R\Hom_{k[x_1,\dots,x_n]}(A,k[x_1,\dots,x_n][n])$

Then $\injdim R_A=\injdim {_A}R=0$, $\Ext^i_A(S,R)=\delta_{i=0}T$ for all simple left $A$-modules $S$.

\begin{defn}
	$R$ is called balanced if 
	\begin{itemize}
		\item $\injdim {_A}R=\injdim R_A=0$
		\item $R\Hom_A(k,R)=k$ and $R\Hom_{A^{op}}(k,R)=k$
	\end{itemize}
\end{defn}

So the idea we wrote down above is very similar to the idea of a balanced dualizing complex (it is stronger).

\subsection{Bass' Theorem}
\begin{thm}[Bass]
	Let $A$ be a local Noetherian algebra with balanced dualizing complex. Let $X\in\Dbperf(\Amod)$ with finite injective dimension. Then 
	\[\injdim X=\sup X+\depth A\]
\end{thm}
\begin{cor}
	Under the same hypotheses, let $M$ be a finitely-generated left $A$-module with finite injective dimension. Then 
	\[\injdim M=\depth A\]
\end{cor}
\begin{rmk}
	Notice that this also implies that the depth of $A$ is finite.
\end{rmk}

We begin the proof of the theorem with a lemma or two:
\begin{lem}
	Suppose $Y\in\Db_{fg}(\Amod)$. Then 
	\[\projdim Y=\sup\{R\Hom_A(Y,k)\}\]
	and 
	\[\sup Y=-\inf\{R\Hom_A(Y,k)\}\]
\end{lem}
\begin{prf}
	To see this, let $P$ be a minimal projective resolution of $Y$. Write $S=\sup(U)$ and $d=\projdim Y$.

	The idea for completing this proof again leverages that $\partial(P^i)\subseteq J(A)P^{i+1}$ for all $i$. Then notice that 
	everything is in the kernel of $\partial^\ast$, the induced map from the functor $\Hom_A(-,k)$, which then implies that
	$\sup(\Hom_A(P,k))=d$ and that $\inf \Hom_A(P,k)=-s$ is clear. Then the result follows from using that $R\Hom_A(Y,k)=\Hom_A(P,k)$.
\end{prf}
Let's get another lemma!
\begin{lem}
	Suppose $X\in\Db_{fg}(A)$. Then 
	\[\injdim\bbD(X)\le \projdim X\]
	and 
	\[\projdim \bbD(X)\le\injdim X.\]
\end{lem}
\begin{prf}
	Recall here that $\bbD(X)=R\Hom_A(X,R)$ and $\bbD^\circ(X_A)=R\Hom_{A^{op}}(X,R).$ Then suppose that $d=\projdim X<\infty.$ Let $P$ be a minimal 
	free resolution of $X$. Write $I_R$ to denote a resolution of $R$ as an $A$-bimodules such that each term is injective on the right. Since $\injdim R=0$, $I_R^{>0}=0.$

	But now $\bbD(X)=R\Hom_A(X,R)=\Hom_A(P,I_R)$. Now note that 
	\[\Hom_A(P,I_R)^i=\oplus_{w\in\bbZ}\Hom(P^w,I_R^{i+w})=\oplus (I_r^{i+w})^{\operatorname{rank}(P^w)}\]
	so in particular, it is injective. Furthermore, $\Hom_A(P,I_k)^i=0$ whenever $i>d.$

	But then $\bbD(X)$ has an injective resolution that is bounded above at $d$.

	But then we know that 
	\[\projdim \bbD(X)=\sup\{R\Hom_{A^{op}}()\bbD(X),k)\}=\sup\{R\Hom_A(\bbD^\circ(k),\bbD^\circ\bbD(X))\}\]
	where the last equality is by the (earlier) duality theorem. But $\bbD^\circ(k)=k$ and $\bbD^\circ\bbD(X)=X$,
	and so since $\sup\{R\Hom_A(k,X)\}\le\injdim X$, we get
	\[\projdim\bbD(X)\le\injdim X\]

	I think I missed the other one?!
\end{prf}
\begin{cor}\label{cor-equalities}
	\[\projdim \bbD(X)=\injdim X\] 
	and 
	\[\injdim X=\sup\{R\Hom_A(k,X)\}\].
\end{cor}
\begin{lem}
	If $R$ is the balanced dualiznig complex of $A$,
	\[\depth A=-\sup(R)\ge 0\]
\end{lem}
\begin{prf}
	\begin{align*}
		\depth A &= \inf\{R\Hom_A(k,A)\}\\
		&= \inf\{R\Hom_{A^{op}}(\bbD(A),\bbD(k))\}\\
		&=\inf\{R\Hom_{A^{op}}(R,k)\}\\
		&= -\sup R.
	\end{align*}
\end{prf}
Next recall some identities from last week:
\begin{lem}
	Let $X\in\Dbperf(\Amod)$. Then if $\Hom(H^s(Y),k)\ne 0$ where $s=\sup Y$, then 
	\[\sup\{R\Hom(X,Y)\}=\projdim X+\sup Y.\]
\end{lem}

Finally we get around to proving the theorem:
\begin{prf}[of Bass' theorem]
	Suppose $\injdim X<\infty.$  Then by corollary~\ref{cor-equalities} above, $\projdim\bbD(X)<\infty$. Then
	\begin{align*}
		\sup X &=\sup\{R\Hom_A(A,X)\}\\
		&=\sup\{R\Hom_{A^{op}}(\bbD(X),\bbD(A))\}\\
		&=\sup\{R\Hom_{A^{op}}(\bbD(X),R)\}\\
		&=\projdim\bbD(X)+\sup R\\
		&=\injdim X-\depth A.
	\end{align*}
	And that's what we wanted to prove!
\end{prf}
\begin{rmk}
	The preceding proof was adapted by Peter Jorgensen's paper. He apparently has very clear exposition 
	and arguments where each step is carefully justified. The paper in question here is entitled something like \textit{Homological Identiies}
	but apparently all of his papers have a similar level of polish and James recommends reading them all.
\end{rmk}

\subsection{No Holes Theorem}
\begin{thm}[No Holes]
	For every $i$ between $\depth M$ and $injdim M$,
	\[\Ext_A^i(k,M)\ne 0.\]
\end{thm}

\section{May 31, 2019}
Today we begin our discussion of Artin-Schelter regular algebras.
\subsection{Some History/Comments}
In the 1980's ('86), Artin and Schelter started a project of classifying connected graded regular algebras of global dimension 3. 
This is somehow comparible to quantum or noncommutative projective planes $q\bbP^2$ or $\bbP^2_{nc}$.

There were actualy some examples of such algebras discovered befor this, however. In 1982, 
Sklyanin found examples of regular algebras of global dimension 4 in the suty of they Yang-Baxter equation and quantum groups.

In 1991, the ``AS'' project was finished by Artin, Schelter, Tate, and Ven den Bergh. All together they constitute about 
150 pages and the result was that there were 14 families. Later on these were named Artin-Shelter regular algebras.

\subsection{Examples}
\begin{ex}
	Let $d=\gldim A$. Then the following are AS-regular:
	\begin{itemize}
		\item When $d=0$, only $k$ is AS regular.
		\item When $d=1$, $k[x]$ is AS regular when $\deg x>0$.
		\item For $d=2$, if $k=\bar k$, then $k_q[x,y]:=k\langle x,y\rangle/(yx-qxy)$ and $k_J[x,y]=k\langle x,y\rangle/(yx-xy-x^2)$ are examples.
	\end{itemize}
\end{ex}
\subsection{Definitions}
Recall that $A$ is called connected graded if $A=\oplus_\bbN A_i$ with $1_A\in A_0=k$ and $A_iA_j\subseteq A_{i+j}$. Then $\bbN$ is called 
an internal (or Adams) grading, which is NOT the complex grading -- in particular the Koszul sign rule does \textbf{not} apply to this grading.
\begin{defn}
	A connected graded algebra $A$ is called \textbf{Artin-Schelter regular} (AS regular) if 
	\begin{itemize}
		\item $\gldim A=d<\infty$
		\item Let $k=A/A_{\ge 1}$ be the trivial module. Then 
	\[\Ext_A^i(k,A)=\Ext_{A^{op}}^i(k,A)=\left\{\begin{array}{lr}0, & i\ne d\\ k,& i=d\end{array}\right.\]
	\end{itemize}
\end{defn}
\begin{rmk}
	Notmally one requires that the algebra also has finite GK dimension (no idea what this is), but we will assume that our algebras are Noetherian, 
	which obviates this restriction.
\end{rmk}
\begin{prop}
	If $A$ is connected graded and commutative, then the following are equivalent:
	\begin{itemize}
		\item $A$ is AS regular.
		\item $A$ has finite global dimension
		\item $A=k[x_1,\cdots,x_n]$.
	\end{itemize}
\end{prop}
\begin{ex}
	A \textbf{non-example} is that $k\langle x_1,\cdots,x_n\rangle$ is \textit{not} AS-regular.
\end{ex}
\begin{rmk}
	A connected graded algebra is a graded version of a complete local algebra. That is, $A=\varprojlim A/A_{\ge n}$.
\end{rmk}
\begin{rmk}\label{rmk-AB}
	\textsc{(Auslander-Buchsbaum formula)}: Suppose $A$ satisfies the $\chi$ condition (that is, for all finitely generated $A$-modules $M$, $\dim\Ext_A^i(k,M)<\infty$ for all $i$)
	If $M$ has finite projective dimension then 
	\[\projdim M=\depth A-\depth M.\]
\end{rmk}

\subsection{Today's Main Result}
\begin{thm}
	Let $A$ be a connected graded Notherian algebra with finite global dimension. Then the following are equivalent:
	\begin{itemize}
		\item $A$ is AS regular
		\item $A$ satrisfies the $\chi$ condition
		\item For any $X\in\Db(\Amod_{fg})$ (can also switch sides) then $\dim\Ext_A^i(k,X)<\infty,\forall i$.
		\item $A$ has a balanced dualizing complex.
	\end{itemize}
\end{thm}
To prove this, we will first need two lemmas.
\begin{lem}
	Suppose $A$ satisfies $\chi$  and $\gldim A=d<\infty$. Then 
	\begin{itemize}
		\item $\depth A=d$
		\item $\projdim k=d$.
	\end{itemize}
\end{lem}
\begin{prf}
	Proving this uses the Auslander-Buchsbaum formula from remark~\ref{rmk-AB}. For any finitely-generated module $M$,
	\[\depth A=\projdim M=\depth M\ge \projdim M\]
	and so by definition, $\depth A\ge \gldim A=d$.

	To prove the other inequality we need to unpack the definition of depth:
	\[\depth A=\inf(R\Hom(k,A))\le \sup(R \Hom(k,A))=\projdim k\le \gldim A=d.\]
	Thus $\depth A=d$. But we also havev $\projdim k$ in there, so the second equality also holds.
\end{prf}
\begin{lem}
	Let $F:\Amod_{fd}\to \modA_{fd}$ and $G:\modA_{fg}\to\Amod_{fg}$ be inverse functors. Then $F({_A}k)=k_A$
	and $G(k_A)={_A}k.$
\end{lem}
\begin{prf}
	This is almost trivial if you think hard about where a simple trivial module has to be sent.
\end{prf}
\begin{prf}[of theorem]
	(1)$\Rightarrow$(3): By definition we know what $\Ext_A^i(k,A)$ is. Define $\calT$ to be the triangulated subcategory of 
	$\Db_{fg}(\Amod)$ of all objects $X$ such that $\dim\Ext_A^i(k,X)<\infty$ for all $i$.

	It is easy to check that this is a triangulated subcategory. By the definition of AS-regularity, $A$ is in $\calT$. And we know $\Db_{fg}(\Amod)=\Dbperf(\Amod)=\langle A\rangle$, so $\calT$ is 
	the entire category $\Db_{fg}(\Amod).$

	(3)$\Rightarrow$(2) is left as an exercise, but we can do it by taking $X=M$.

	(2)$\Rightarrow$(1) By the first lemma, we have $\Ext_A^i(k,A)=0$ when $i\ne d$ and is finite dimensional otherwise.
	Let $F(-)=\Ext_A^d(-,A)$ and $G(-)=\Ext^d_{A^{op}}(-,A)$. Now note that 
	\[F(M)=R\Hom_A(M,A)[-d]=R\Hom_A(M,A[d])\]
	and let $R=A[d]$. The claim is that $R$ is a dualizing complex. This implies that $F(M)=\bbD(M)$ and $G(N)=\bbD^\circ(N)$,
	and we can apply the lemma to show that $\Ext_A^d(k,A)=F(k)=k$ and $\Ext_{A^{op}}^d(k,A)=k$, so we're done.
\end{prf}
